Name matching through deep learning

### 2018-11-16 old/paper2018/introduction2.tex
reorgonized directory
added mudgal ref

Merging datasets is a key operation for data analytics.  A frequent
requirement for merging is joining across columns that have
different surface forms for the same entity.  For instance, the name of a person might be represented as \textit{Douglas Adams}, \textit{Douglas Noel Adams}, \textit{D. Adams} or \textit{Adams, Douglas}.  Similarly, ontology alignment can require recognizing distinct surface forms of
the same entity, especially when ontologies are independently
developed.  This problem occurs for many entity types such as people's names, company names, addresses, product descriptions, conference venues, or even people's faces.  Data management systems have however, largely focussed solely on equi-joins, where string or numeric equality determines which rows should be joined, because such joins are efficient.

We propose a different approach to joining different surface representations of the same entity, inspired by recent advances in deep learning.  Our approach depends on (a) mapping surface forms into sets of vectors such that forms for the same entity are closest in vector space, (b) indexing these vectors to find the forms that can be potentially joined together.  The approach is general, in the sense that once a model has been built for a specific semantic type (e.g. people, companies or faces) it can be used for joining any two datasets which share that semantic type.  It is also efficient because indexing uses space partitioning algorithms (such as approximate nearest neighbor) to find surface forms that are potentially joinable, thus eliminating large parts of the vector space from consideration.  Further, nearest neighbor algorithms have been applied to billions of vectors \cite{JDH17}, so the approach is practical for most datasets.

To test the feasibility of these ideas, we used Wikidata as ground truth to build models for datasets with 1.1M people's names (about 200K identities) and 130K company names (70K identities).  The problem of mapping vectors for the same entity closer in vector space than vectors for other entities is known in the literature as deep metric learning.  Deep metric learning is known to be a difficult problem as studied in the space of face recognition and person-re-identification \cite{DBLP:conf/cvpr/SchroffKP15}, \cite{8445716}, \cite{DBLP:journals/corr/abs-1802-03170}.  As a result, there is a significant amount of research on two aspects of training these networks: (a) how to choose samples for efficient learning, and (b) what constitutes a good loss function.  In building models for entity names, we had to adapt these techniques for triplet selection and loss functions because matching entity names has different characteristics, as we describe below.

@@

We also investigated the effect of using multiple local loss functions which have been proposed in the literature for improving deep metric learning, e.g., the triplet loss \cite{DBLP:conf/cvpr/SchroffKP15}, improved loss \cite{Zhang:2016:DML:3088616.3088665}, and angular loss \cite{DBLP:journals/corr/abs-1708-01682} functions.  For the problem of building models for datasets dominated by hard negatives, we found that an adaptation to the triplet loss function proposed in \cite{DBLP:conf/cvpr/SchroffKP15} was better than three other functions that have been used in the literature. All code and models for this paper are at \url{https://github.com/yehudagale/fuzzyjoiner} and \url{https://drive.google.com/open?id=1zivCTGkq2_AkfjGLHMnlehzTmYUwcQ9e} respectively.
%
%\begin{table}[!htb]
%    \caption{Example of a merge problem}
%    \begin{subtable}{.5\linewidth}
%      \centering
%        \caption{Table A}
%        \begin{tabular}{|l|l|}
%          \hline
%           Dept & Name \\
%           \hline
%           1    & Douglas Adams \\
%           2    & John Adams \\
%           3  & Douglas Bard \\
%           \hline
%        \end{tabular}
%    \end{subtable}%
%    \begin{subtable}{.5\linewidth}
%      \centering
%        \caption{Table B}
%        \begin{tabular}{|l|l|}
%          \hline
%           Name & Salary \\
%           \hline
%           Douglas Noel Adams & 2000 \\
%           Adams, John & 3000 \\
%           John Smith & 3000 \\
%           \hline
%        \end{tabular}
%    \end{subtable}
%    \label{table-example}
% \end{table}

>>>>

Merging datasets frequently requires joining across columns that have
different surface forms for the same entity.  For instance, the name of a person might be represented as \textit{Douglas Adams}, \textit{Douglas Noel Adams}, \textit{D. Adams} or \textit{Adams, Douglas}.  Similarly, ontology alignment can require recognizing distinct surface forms of
the same entity, especially when ontologies are independently
developed.  This problem occurs for many entity types such as people's names, company names, addresses, product descriptions, conference venues, or even people's faces.  Data management systems have however, largely focussed solely on equi-joins, where string or numeric equality determines which rows should be joined, because such joins are efficient.

We propose a different approach to joining different surface representations of the same entity, inspired by recent advances in deep learning.  Our approach depends on (a) mapping surface forms into sets of vectors such that forms for the same entity are closest in vector space, (b) indexing these vectors to find the forms that can be potentially joined together.  The approach is general, in the sense that once a model has been built for a specific semantic type (e.g. people, companies or faces) it can be used for joining any two datasets which share that semantic type, provided of course that the sample characteristics match.  It is also efficient because indexing uses space partitioning algorithms (such as approximate nearest neighbor) to find surface forms that are potentially joinable, thus eliminating large parts of the vector space from consideration.  Further, nearest neighbor algorithms have been applied to billions of vectors \cite{JDH17}, so the approach is practical for most datasets.

To test the feasibility of these ideas, we used Wikidata as ground truth to build models for datasets with 1.1M people's names (about 200K identities) and 130K company names (70K identities).  The problem of mapping vectors for the same entity closer in vector space than vectors for other entities is known in the literature as deep metric learning.  Deep metric learning is known to be a difficult problem as studied in the space of face recognition and person-re-identification \cite{DBLP:conf/cvpr/SchroffKP15}.  As a result, there is a significant amount of research on two aspects of training these networks: (a) how to choose samples for efficient learning, and (b) what constitutes a good loss function.  In building models for entity names, we had to adapt these techniques for triplet selection and loss functions because matching entity names has different characteristics, as we describe below.

@@

We also investigated the effect of using multiple local loss functions which have been proposed in the literature for improving deep metric learning, e.g., the triplet loss \cite{DBLP:conf/cvpr/SchroffKP15}, improved loss \cite{Zhang:2016:DML:3088616.3088665}, and angular loss \cite{DBLP:journals/corr/abs-1708-01682} functions.  For the problem of building models for datasets dominated by hard negatives, we found that an adaptation to the triplet loss function proposed in \cite{DBLP:conf/cvpr/SchroffKP15} was better than three other functions that have been used in the literature. 
%
%\begin{table}[!htb]
%    \caption{Example of a merge problem}
%    \begin{subtable}{.5\linewidth}
%      \centering
%        \caption{Table A}
%        \begin{tabular}{|l|l|}
%          \hline
%           Dept & Name \\
%           \hline
%           1    & Douglas Adams \\
%           2    & John Adams \\
%           3  & Douglas Bard \\
%           \hline
%        \end{tabular}
%    \end{subtable}%
%    \begin{subtable}{.5\linewidth}
%      \centering
%        \caption{Table B}
%        \begin{tabular}{|l|l|}
%          \hline
%           Name & Salary \\
%           \hline
%           Douglas Noel Adams & 2000 \\
%           Adams, John & 3000 \\
%           John Smith & 3000 \\
%           \hline
%        \end{tabular}
%    \end{subtable}
%    \label{table-example}
% \end{table}

### 2018-11-16 old/paper2018/relatedWork.tex
added mudgal ref

The idea of building joint embeddings for merging datasets followed by nearest neighbors search has been applied recently to the problem of linking relational tuple embeddings with embeddings of other relational tuples or unstructured text \cite{Bordawekar18}, \cite{IDEL18}).  For the problem of linking tuples, each model that is learnt is specific to the database it was trained on.  Our focus is on techniques that can be used to develop general purpose embedding models for merging alternate surface forms of key entities.  Once such models are built, they can be applied to joining any two datasets that share that semantic type.

Metric learning is a well studied problem in the face recognition literature, e.g., \cite{DBLP:conf/cvpr/SchroffKP15}, with a rich literature in triplet mining techniques.  The closest approach to ours is the use of nearest neighbors algorithms for semi-hard triplet mining \cite{DBLP:journals/corr/KumarHC0D17}.  For semi-hard triplet mining, one cannot look at fixed neighborhood sizes in building triplets.  If all the positives are further away from the anchor than the negatives in a given neighborhood size of $k$, it means that $k$ needs to be expanded until a neighborhood size is found that has the right characteristics.  Our approach in including `hard negatives' means we can use a fixed $k$ to generate samples.  An additional benefit is that at least for certain types of datasets, we show that metric learning with hard negatives is more effective than semi-hard mining.

>>>>

The idea of building joint embeddings has been applied recently to the problem of linking relational tuples for entity resolution \cite{Mudgal},\cite{Bordawekar18}).  For the problem of linking tuples, each model that is learnt is specific to the database it was trained on because the attributes of an entity vary significantly across databases.  Our focus is on techniques for name matching, and we develop general purpose embedding models for merging alternate surface forms of key entities.

Metric learning is a well studied problem in face recognition, e.g., \cite{DBLP:conf/cvpr/SchroffKP15}, with a rich literature in triplet mining techniques.  The closest approach to ours is the use of nearest neighbors algorithms for semi-hard triplet mining \cite{DBLP:journals/corr/KumarHC0D17}.  For semi-hard triplet mining, one cannot look at fixed neighborhood sizes in building triplets.  If all the positives are further away from the anchor than the negatives in a given neighborhood size of $k$, it means that $k$ needs to be expanded until a neighborhood size is found that has the right characteristics.  Our approach in including `hard negatives' means we can use a fixed $k$ to generate samples.  An additional benefit is that at least for certain types of datasets, we show that metric learning with hard negatives is more effective than semi-hard mining.

### 2018-11-16 old/paper2018/samplepaper.tex
merge
figure hacking
added mudgal ref

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage[caption=false]{subfig}
\usepackage{multirow}

@@

\begin{document}
%
\title{Name matching through deep learning}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Kavitha Srinivas\inst{1} \and
Yehuda Gale\inst{2} \and
Julian Dolby\inst{1}}
%
\authorrunning{K. Srinivas et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{IBM Research \and
Rutgers University}

%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Merging datasets is a key operation for data analytics.  A frequent
requirement for merging is joining across columns that have
different surface forms for the same entity (e.g., the name of a
person might be represented as \textit{Douglas Adams} or
\textit{Adams, Douglas}).  Similarly,
ontology alignment can require recognizing distinct surface forms of
the same entity, especially when ontologies are independently
developed.  However, data management systems are currently limited
to performing merges based on string equality, or at best using
string similarity.  We propose an approach to performing merges
based on deep learning models.  Our approach depends on (a) creating
a deep learning model that maps surface forms of an entity into a
set of vectors such that alternate forms for the same entity are
closest in vector space, (b) indexing these vectors using a nearest
neighbors algorithm to find the forms that can be potentially joined
together.  To build these models, we had to adapt techniques from
metric learning due to the characteristics of the data; specifically we describe 
novel sample selection techniques and loss functions that work for this problem.  
To evaluate our approach, we used Wikidata as ground truth
and built models from datasets with approximately 1.1M people's names
(200K identities) and 130K company names (70K identities).  We developed models that 
allow for joins with precision@1 of .75-.81 and
recall of .74-.81.  We make the models available for aligning people or companies across multiple datasets.

@@

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{8}
\bibitem{ref_article1}
Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)

\bibitem{ref_lncs1}
Author, F., Author, S.: Title of a proceedings paper. In: Editor,
F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
Springer, Heidelberg (2016). \doi{10.10007/1234567890}

\bibitem{ref_book1}
Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
Location (1999)

\bibitem{ref_proc1}
Author, A.-B.: Contribution title. In: 9th International Proceedings
on Proceedings, pp. 1--2. Publisher, Location (2010)

\bibitem{ref_url1}
LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
Oct 2017
\end{thebibliography}
\end{document}

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage[caption=false]{subfig}
\usepackage{multirow}
\usepackage{wrapfig}

@@

\begin{document}
%
\title{Name matching through deep learning}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Kavitha Srinivas\inst{1} \and
Abraham Gale\inst{2} \and
Julian Dolby\inst{1}}
%
\authorrunning{K. Srinivas et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{IBM Research \and
Rutgers University}

%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Merging datasets frequently requires joining across columns that have
different surface forms for the same entity (e.g., the name of a
person might be represented as \textit{Douglas Adams} or
\textit{Adams, Douglas}).  However, data management systems are currently limited
to performing merges based on string equality, or at best using
string similarity.  We propose an approach to name matching based on deep learning models.  
Our approach depends on (a) creating
a deep learning model that maps alternate forms for the same entity into vector space such that alternate forms are closest, (b) indexing these vectors using a nearest
neighbors algorithm to find the forms that can be potentially joined
together.  To build these models, we had to adapt techniques from
metric learning due to the characteristics of the data; specifically we describe 
novel sample selection techniques and loss functions that work for this problem.  
To evaluate our approach, we used Wikidata as ground truth
and built models from datasets with approximately 1.1M people's names
(200K identities) and 130K company names (70K identities).  We developed models that 
allow for joins with precision@1 of .75-.81 and
recall of .74-.81.  We make the models available for aligning people or companies across multiple datasets.

@@

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{paper}

\end{document}
